---
title: "my-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(koo)
```


Koo is an R package which simplifies and enhances Box-Cox transformations. This package provides a suite of functions which allow a user to apply, analyze and visualize Box-Cox transformations. Six functions are included in the package.

1. `box_transform()`: For applying the Box-Cox transformation

2. `inverse_box_transform()`: For transforming a Box-Cox transformed data set back to its original

3. `BCVisual()`: For visualizing a plot of the transformed data in comparison to the original data.

4. `optimallambda()`: For estimating the optimal lambda value for a given data set. Allows for most optimal transformation

5. `box_diagnostic_plot()`: For visualizing the normal diagnostic plots of the transformed data and comparing this with the original diagnostic plots.

6. `diagnostic_sim()`: For generating a diagnostic summary of transformed data over a range of different lambda values.



