---
title: "my_vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(koo)
```

Koo is an R package which simplifies and enhances Box-Cox transformations. This package provides a suite of functions which allow a user to apply, analyze and visualize Box-Cox transformations. Six functions are included in the package.

1. `box_transform()`: For applying the Box-Cox transformation

2. `inverse_boxcox()`: For transforming a Box-Cox transformed data set back to its original

3. `box_plot`: For visualizing a plot of the transformed data in comparison to the original data.

4. `estimate_optimal_lambda()`: For estimating the optimal lambda value for a given data set. Allows for most optimal transformation

5. `box_diagnostics()`: For visualizing the normal diagnostic plots of the transformed data and comparing this with the original diagnostic plots.

6. `diagnostic_sim()`: For generating a diagnostic summary of transformed data over a range of different lambda values.

___koo___ will run in Windows, Mac OS X, or Linux. To install it you first need to install [R](https://cran.r-project.org/). Installing [RStudio](https://www.rstudio.com/) as a desktop environment for using R is also recommended for ease of use.

Type into the R command prompt
```{r,eval=FALSE}
install.packages("devtools")
devtools::install_github('Killian-s03/koo')
```

To install the latest development from the ___koo___  [GitHub page](https://github.com/Killian-s03/koo). 

To load the package into R type:

```{r,eval=FALSE}
library(koo)
```


This will load all functions within the ___koo___ package.


##'inverse_boxcox'

The inverse_boxcox function applies the inverse Box-Cox transformation to a dataset that has been previously transformed. This method restores the original scale of the data using the specified $\lambda$.

```{r}
transformed <- c(2.3, 1.8, 3.1)
lambda <- 0.5
inverse <- inverse_boxcox(transformed, lambda)
print(inverse)
```

##'estimate_optimal_lambda

The estimate_optimal_lambda function provides a flexible way to determine the optimal value of $\lambda$ for the Box-Cox transformation. This function can handle a variety of input types, including data frames, matrices, and numeric vectors, and supports user-defined ranges and resolution for $\lambda$ values.
It works by evaluating the log-likelihood of a fitted linear model for various lambda values and selecting the one that maximizes the log-likelihood.

```{r}
set.seed(2000)
mat <- cbind(
  y = rexp(100, rate = 0.3),
  x1 = runif(100, 2, 8),
  x2 = runif(100, 10, 20)
)
lambda_optimal <- estimate_optimal_lambda(mat, lambda_range = c(-2, 2), resolution = 100)
print(lambda_optimal)
```


